### 2012-06-04 paper/referee_response.txt
response to all apart from one referee comment

excellent suggestions, improve the paper. Respond inline

	--Sec 2.2: "current" generation of large survey imagers would be more
	  accurate than "next" generation.

Done.

	--Sec 3.1: "We We"

Done.

	--Sec 5.1: please include some description of the density and/or total
	  number of sources used, either here or in Sec. 2.2.  Also insert "16
	  deg^2" before "patch of sky".

We explicitly put the source density into the last sentence of 2.1 and we re-point the reader to the parameter table in the last sentence prior to Section 5.1. We included the solid angle of the total patch, as requested.

	--suggestion for going a bit deeper in Sec 5.1: some surveys may find
	  it difficult to rotate their hardware randomly with every shot.
	  Could you do a survey E which has random pointing centers but no
	  rotation?

The referee is absolutely right that there are cameras that cannot be rotated with respect to the sky. That is one of the reasons why we designed Survey C without rotation. Since Survey C calibrates just as well as Survey D, we conclude that rotation is not necessary for good self-calibration. We agree that a Survey E would confirm that, but we believe that Survey C already makes this point well. We added a few sentences to the discussion of Survey C and D about this point in Sections 5.1 and 5.2.

	--suggestion for going a bit deeper in Sec 5.2 (and/or Sec 4): you
	  show the residuals in the instrument response, but more relevant
	  might be the residuals in the final survey.  For Survey A (and to a
	  large extent B), the residuals in the final survey are just as bad
	  as in the instrument response.  But in Surveys C and D, they get beat
	  down because the calibration error on a given piece of sky changes
	  from observation to observation.  It would be instructive to add a
	  performance metric which includes this effect.

In Addition to these changes, we also made the following small changes:
	- MeasuredCatalog.true_invvar - multiplied by flat-field one time too many??
	- Generate epsilon now per measurement and not per source

Neither made significant changes to content of papers