### 2012-07-27 paper/referee_response.txt
made all referee changes

The review makes some excellent suggestions to improve the paper. Please see the inline responses below:

	--Sec 2.2: "current" generation of large survey imagers would be more
	  accurate than "next" generation.

Done.

	--Sec 3.1: "We We"

Done.

	--Sec 5.1: please include some description of the density and/or total
	  number of sources used, either here or in Sec. 2.2.  Also insert "16
	  deg^2" before "patch of sky".

We explicitly put the source density into the last sentence of 2.1 and we re-point the reader to the parameter table in the last sentence prior to Section 5.1. We included the solid angle of the total patch, as requested.

	--suggestion for going a bit deeper in Sec 5.1: some surveys may find
	  it difficult to rotate their hardware randomly with every shot.
	  Could you do a survey E which has random pointing centers but no
	  rotation?

The referee is absolutely right that there are cameras that cannot be rotated with respect to the sky. That is one of the reasons why we designed Survey C without rotation. Since Survey C calibrates just as well as Survey D, we conclude that rotation is not necessary for good self-calibration. We agree that a Survey E would confirm that, but we believe that Survey C already makes this point well. We added a few sentences to the discussion of Survey C and D about this point in Sections 5.1 and 5.2.

	--suggestion for going a bit deeper in Sec 5.2 (and/or Sec 4): you
	  show the residuals in the instrument response, but more relevant
	  might be the residuals in the final survey.  For Survey A (and to a
	  large extent B), the residuals in the final survey are just as bad
	  as in the instrument response.  But in Surveys C and D, they get beat
	  down because the calibration error on a given piece of sky changes
	  from observation to observation.  It would be instructive to add a
	  performance metric which includes this effect.

This is an excellent suggestion. We have added Figures 7 to 11, which show the residuals in the final survey.

In Addition to these changes, we also made the following small changes:
	- We removed "Large-Scale" from the paper title
	- We corrected a few minor bugs in the code, none of which made made significant changes to the contents of the paper.
	- We have modified the form of the true instrument response (last equation in 2.2.3) to now include a (1+a+b)**-0.75 factor. This reduces the magnitude of the high frequency sine and cosine contributions. We did this so that the best-in-basis instrument response was closer to the true instrument response. 
	- The code used in this work is publicly available and we added the github links in Section 2 and the Acknowledgements 
	- We re-ran the simulations with a higher source density (d = 300 not d = 100 deg^-2), as this makes the survey residual plots (Figure 7 - 11) clearer.
	- For clarity, we have slightly reworded the abstract.
	- Finally, we have also corrected a few typos.