### 2012-08-20 whitepaper/pixel_level.tex
start

\documentclass[12pt]{article}
\begin{document}
Hello world.
\end{document}