### 2012-08-20 whitepaper/pixel_level.tex
pixel signal models

\documentclass[12pt,preprint,dvips]{aastex}

\newcommand{\foreign}[1]{\textsl{#1}}
\newcommand{\etal}{\foreign{et~al.}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\SDSS}{\project{SDSS}}

\newcommand{\expectation}[1]{\tilde{#1}}
\newcommand{\sky}{^{(\mathrm{sky})}}

\begin{document}

\section*{Can we infer the pixel-level flat-field without an internal calibration source?}

\noindent
David W. Hogg (NYU)
\\
\textsl{Max-Planck-Institut f\"ur Astronomie, Heidelberg, Germany}

We have shown with large optical imaging surveys that the redundant
information coming from the multiple observations of stars is
sufficient to infer the large-scale sensitivity (effective area or
flat-field or vignetting) of an astronomical imaging device
(\citealt{ubercal}, \citealt{schlafly}, \citealt{holmes}).  In these
self-calibration experiments, the sensitivity function has had only
hundreds or thousands of parameters.  We have not attempted to obtain
millions of pixel-level parameters from the data and we have not
considered arbitrarily non-smooth two-dimensional sensitivity
functions.

With observations of billions of sources on the sky, however, it might
be possible to infer millions of sensitivity parameters, even as many
or more sensitivity parameters than there are pixels in the imaging
device.  After all, we know a lot about the sky, many sources are
observed multiple times, and every source touches multiple pixels even
in a single visit.

In detail, the sensitivity of a pixel can be a function of spectral
energy distribution (different pixels will be differently sensitive to
sources with different colors or spectra) and a function of
intra-pixel illumination pattern (the pixel might have a more
sensitive region and some less sensitive regions).  Also, diffuse
illumination, such as uniform sky background, might illuminate the
pixel from a different mixture of angles than compact illumination,
such as from a star, so there might be, effectively, different pixel
sensitivities for the sky and for sources.  Indeed, reflections in the
\SDSS\ camera led to problems of this kind, which invalidated the
sky-based flat-fields (\citealt{ubercal}).  For all these reasons we
might actually have many sensitivity parameters per pixel; this makes
our task harder quantitatvely, but perhaps not qualitatively.

\paragraph{varying levels of dumbness:}
There are various thought experiments or considerations that suggest
that a pixel-level self-calibration is possible.  One observation is
that if a pixel is unusually low in sensitivity, it's mean output
signal will be low on average for the whole data set.  This statement
could be made more precise for the sky flat by considering only
observations in which the pixel is not near any detectable source.  It
could be made even more precise if each observation is divided by the
mean pixel value, to account for variations in sky brightness.  These
kinds of operations or considerations could lead to a very good sky
flat for any device, and is not unlike what is usually done.

As we said, however, each pixel might have a different response to
compact sources than it does to the diffuse sky, so slightly less-dumb
methods might be called for.  For example, consider all overlapping
$3\times 3$-pixel patches in a full data set.  Most of these patches
would be far from any detectable source and could be labeled ``sky''.
A fraction would be near or on top of compact, detected sources and
could be labeled ``source''.  This labeling could be
non-comprehensive; there could be intermediate patches labeled
``neither''; this is just an engineering consideration.  Less dumb
still would be to make bigger patches (say $5\times 5$ or larger) or
to make use of the residuals away from a physically motivated model of
the observed images.

\paragraph{pixel signal models}
Sticking with the intermediate level of dumbness for now: In a set of
$3\times 3$ sky patches, the expected signal
$\expectation{D}_{ij}$---the expected number of electrons or data
numbers read out---in pixel $i$ of exposure $j$ might be a combination
of zero-level, dark-current, and sky-photon signals
\begin{equation}
\expectation{D}_{ij} = z_i + d_i\,t_j + f\sky_i\,S\sky_j\,t_j
\end{equation}
where $i$ labels the device pixel ($i$ maps back to a pixel location
in the original imaging device), $j$ labels the exposure, each of
which might have a slightly different sky level, $z_i$ is a zero-level
or bias for pixel $i$, $d_i$ is a dark current rate or signal per time
in the absence of illumination, $t_j$ is the exposure time for
exposure $j$, $f\sky_i$ is the sensitivity of pixel $i$ to diffuse sky
illumination, and $S\sky_j$ is the intensity of the sky in exposure
$j$ (interpolated to pixel $i$ if there are gradients).

In a $3\times 3$ source patch, the expectation is similar...
\begin{equation}
\expectation{D}_{ij} = z_i + d_i\,t_j + f\sky_i\,S\sky_j\,t_j
\end{equation}

\begin{thebibliography}{70}
\bibitem[Holmes \etal(2012)]{holmes}
Holmes,~R., Hogg,~D.~W., \& Rix,~H.-W., 2012, arXiv:1203.6255 
\bibitem[Padmanabhan \etal(2008)]{ubercal}
Padmanabhan,~N., Schlegel,~D.~J., Finkbeiner,~D.~P., \etal, 2008, \apj, 674, 1217
\bibitem[Schlafly \etal(2012)]{schlafly}
Schlafly,~E.~F., Finkbeiner,~D.~P., Juric,~M., \etal, 2012, arXiv:1201.2208 
\end{thebibliography}

\end{document}